---
title: "Biostats_HW2_P6"
author: "Abhishek Ajay (aa4266)"
date: "September 27, 2018"
output:
  html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)

```
#Problem 3

We are given that 8% percent of men and 30% of women are **colorblind**.

**(A)** 

If 10 men were randomly selected for a study on traffic signal perceptions, find the probability that exactly half of them cannot distinguish between red and green. 

So here we have a binomial distribution with the probability of success being the probability that the man selected in colorblind = p = 8% = 0.08

N is the number of trials = 10

x is the number of success = 5

Therefore, X - Random variable denoting colorblind men: *X ~ Bin(N,p)* = *X ~ Bin(10, 0.08)*

Using these values in the dbinom function in R gives us:  P(X = 5) = **10C5 x ((0.08)^5) x ((1-0.08)^(10-5))** = `r dbinom(5, 10, 0.08)`


**(B)** 

What if we would randomly select 10 women instead? How would the probability change? 

So here we have a binomial distribution with the probability of success being the probability that the woman selected in colorblind = p = 30% = 0.3

N is the number of trials = 10

x is the number of success = 5

Therefore, X - Random variable denoting colorblind women: *X ~ Bin(N,p)* = *X ~ Bin(10, 0.30)*

Using these values in the dbinom function in R gives us:  P(X = 5) = **10C5 x ((0.30)^5) x ((1-0.3)^(10-5))**  = `r dbinom(5, 10, 0.30)`

We see that the probability of success increases in the case of women and hence it is more likely to pick up a colorblind woman from a random set of 10 women than a colorlind man from a random set of 10 men.



#Problem 6

Here we import the data set
```{r data_import migraine}

migraine <- read_excel("C:/Users/abhis/OneDrive/Desktop/School/Semester 1/P8130_biostats/HW2/migraine.xlsx")
View(migraine)

```

Here we work on the first, **A** part of the question. Summarizing NDDIE, CSED, Cognitive Evaluation with the original scores.
We take into consideration the follwoing parameters: 

* Minimum value

* 1st Quartile

* Median

* Mean = (Summation of i=1 to n) Xi/ní±–

* 3rd Quartile

* Maximum value

* Number of NA's

* Standard Deviation = Variance^(1/2) 

* Variance = 1/(n-1) * ((Summation of i=1 to n) (Xi - Sample mean)^2)

* Size

```{r summary}


#NDDIE
s_nddie_1 = summary(migraine$NDDIE[which(migraine$Migraine == 1)])

sd_nddie_1 = sd(migraine$NDDIE[which(migraine$Migraine == 1)], na.rm = TRUE)

var_nddie_1 = var(migraine$NDDIE[which(migraine$Migraine == 1)], na.rm = TRUE)

s_nddie_0 = summary(migraine$NDDIE[which(migraine$Migraine == 0)])

sd_nddie_0 = sd(migraine$NDDIE[which(migraine$Migraine == 0)], na.rm = TRUE)

var_nddie_0 = var(migraine$NDDIE[which(migraine$Migraine == 0)], na.rm = TRUE)

with_migraine = data.frame(rbind((data.frame(unclass(s_nddie_1))), 'sd' = sd_nddie_1, 'var' = var_nddie_1, 'size' = nrow(migraine[which(migraine$Migraine == 1),])))

colnames(with_migraine) = 'nddie_w_migraine'

without_migraine = data.frame(rbind((data.frame(unclass(s_nddie_0))), 'sd' = sd_nddie_0, 'var' = var_nddie_0, 'size' = nrow(migraine[which(migraine$Migraine == 0),])))

colnames(without_migraine) = 'nddie_w/o_migraine'

data.frame(with_migraine, without_migraine)

#CSED
s_csed_1 = summary(migraine$CESD[which(migraine$Migraine == 1)])

sd_csed_1 = sd(migraine$CESD[which(migraine$Migraine == 1)], na.rm = TRUE)

var_csed_1 = var(migraine$CESD[which(migraine$Migraine == 1)], na.rm = TRUE)

s_csed_0 = summary(migraine$CESD[which(migraine$Migraine == 0)])

sd_csed_0 = sd(migraine$CESD[which(migraine$Migraine == 0)], na.rm = TRUE)

var_csed_0 = var(migraine$CESD[which(migraine$Migraine == 0)], na.rm = TRUE)

with_migraine_csed = data.frame(rbind((data.frame(unclass(s_csed_1))), 'sd' = sd_csed_1, 'var' = var_csed_1, 'size' = nrow(migraine[which(migraine$Migraine == 1),])))

colnames(with_migraine_csed) = 'csed_w_migraine'

without_migraine_csed = data.frame(rbind((data.frame(unclass(s_csed_0))), 'sd' = sd_csed_0, 'var' = var_csed_0, 'size' = nrow(migraine[which(migraine$Migraine == 0),])))

colnames(without_migraine_csed) = 'csed_w/o_migraine'

data.frame(with_migraine_csed, without_migraine_csed)

#Cognitive Eval

s_mem_1 = summary(migraine$`ABNAS memory`[which(migraine$Migraine == 1)])

sd_mem_1 = sd(migraine$`ABNAS memory`[which(migraine$Migraine == 1)], na.rm = TRUE)

var_mem_1 = var(migraine$`ABNAS memory`[which(migraine$Migraine == 1)], na.rm = TRUE)

s_lang_1 = summary(migraine$`ABNAS language`[which(migraine$Migraine == 1)])

sd_lang_1 = sd(migraine$`ABNAS language`[which(migraine$Migraine == 1)], na.rm = TRUE)

var_lang_1 = var(migraine$`ABNAS language`[which(migraine$Migraine == 1)], na.rm = TRUE)

s_mem_0 = summary(migraine$`ABNAS memory`[which(migraine$Migraine == 0)])

sd_mem_0 = sd(migraine$`ABNAS memory`[which(migraine$Migraine == 0)], na.rm = TRUE)

var_mem_0 = var(migraine$`ABNAS memory`[which(migraine$Migraine == 0)], na.rm = TRUE)

s_lang_0 = summary(migraine$`ABNAS language`[which(migraine$Migraine == 0)])

sd_lang_0 = sd(migraine$`ABNAS language`[which(migraine$Migraine == 0)], na.rm = TRUE)

var_lang_0 = var(migraine$`ABNAS language`[which(migraine$Migraine == 0)], na.rm = TRUE)

with_migraine_mem = data.frame(rbind((data.frame(unclass(s_mem_1))), 'sd' = sd_mem_1, 'var' = var_mem_1, 'size' = nrow(migraine[which(migraine$Migraine == 1),])))

colnames(with_migraine_mem) = 'mem_w_migraine'

without_migraine_mem = data.frame(rbind((data.frame(unclass(s_mem_0))), 'sd' = sd_mem_0, 'var' = var_mem_0, 'size' = nrow(migraine[which(migraine$Migraine == 0),])))
colnames(without_migraine_mem) = 'mem_w/o_migraine'

with_migraine_lang = data.frame(rbind((data.frame(unclass(s_lang_1))), 'sd' = sd_lang_1, 'var' = var_lang_1, 'size' = nrow(migraine[which(migraine$Migraine == 1),])))

colnames(with_migraine_lang) = 'lang_w_migraine'

without_migraine_lang = data.frame(rbind((data.frame(unclass(s_lang_0))), 'sd' = sd_lang_0, 'var' = var_lang_0, 'size' = nrow(migraine[which(migraine$Migraine == 0),])))
colnames(without_migraine_lang) = 'lang_w/o_migraine'

data.frame(with_migraine_mem, with_migraine_lang, without_migraine_mem, without_migraine_lang)

```

##Here we take into consideration the following cutoff for NDDIE and CSED

* NDDIE >= 16

* CSED >= 16

For summarzing the NDDIE and CSED.

```{r summary_cutoff}

#NDDIE

migraine_nddie_c = data.frame(migraine[which(migraine$NDDIE >= 16),])

s_nddie_1_c = summary(migraine_nddie_c$NDDIE[which(migraine_nddie_c$Migraine == 1)])

sd_nddie_1_c = sd(migraine_nddie_c$NDDIE[which(migraine_nddie_c$Migraine == 1)], na.rm = TRUE)

var_nddie_1_c = var(migraine_nddie_c$NDDIE[which(migraine_nddie_c$Migraine == 1)], na.rm = TRUE)

s_nddie_0_c = summary(migraine_nddie_c$NDDIE[which(migraine_nddie_c$Migraine == 0)])

sd_nddie_0_c = sd(migraine_nddie_c$NDDIE[which(migraine_nddie_c$Migraine == 0)], na.rm = TRUE)

var_nddie_0_c = var(migraine_nddie_c$NDDIE[which(migraine_nddie_c$Migraine == 0)], na.rm = TRUE)

with_migraine_c = data.frame(rbind((data.frame(unclass(s_nddie_1_c))), 'sd' = sd_nddie_1_c, 'var' = var_nddie_1_c, 'size' = nrow(migraine_nddie_c[which(migraine_nddie_c$Migraine == 1),])))

colnames(with_migraine_c) = 'nddie_w_migraine_c'

without_migraine_c = data.frame(rbind((data.frame(unclass(s_nddie_0_c))), 'sd' = sd_nddie_0_c, 'var' = var_nddie_0_c, 'size' = nrow(migraine_nddie_c[which(migraine_nddie_c$Migraine == 0),])))

colnames(without_migraine_c) = 'nddie_w/o_migraine_c'

data.frame(with_migraine_c, without_migraine_c)

#CSED

migraine_csed_c = data.frame(migraine[which(migraine$CESD >= 16),])

s_csed_1_c = summary(migraine_csed_c$CESD[which(migraine_csed_c$Migraine == 1)])

sd_csed_1_c = sd(migraine_csed_c$CESD[which(migraine_csed_c$Migraine == 1)], na.rm = TRUE)

var_csed_1_c = var(migraine_csed_c$CESD[which(migraine_csed_c$Migraine == 1)], na.rm = TRUE)

s_csed_0_c = summary(migraine_csed_c$CESD[which(migraine_csed_c$Migraine == 0)])

sd_csed_0_c = sd(migraine_csed_c$CESD[which(migraine_csed_c$Migraine == 0)], na.rm = TRUE)

var_csed_0_c = var(migraine_csed_c$CESD[which(migraine_csed_c$Migraine == 0)], na.rm = TRUE)

with_migraine_csed_c = data.frame(rbind((data.frame(unclass(s_csed_1_c))), 'sd' = sd_csed_1_c, 'var' = var_csed_1_c, 'size' = nrow(migraine_csed_c[which(migraine_csed_c$Migraine == 1),])))

colnames(with_migraine_csed_c) = 'csed_w_migraine_c'

without_migraine_csed_c = data.frame(rbind((data.frame(unclass(s_csed_0_c))), 'sd' = sd_csed_0_c, 'var' = var_csed_0_c, 'size' = nrow(migraine_csed_c[which(migraine_csed_c$Migraine == 0),])))

colnames(without_migraine_csed_c) = 'csed_w/o_migraine_c'

data.frame(with_migraine_csed_c, without_migraine_csed_c)

```
Here we work on the **B** part of the question. We make histograms and boxplots considering the fact that our variables are continuous in nature in the given range.

At first we make histograms in a 1, 2 matrix format for with migraine and wwithout migraine pe variable, i.e NDDIE, CSED, ABNAS Memory, ABNAS Language.

```{r histogram}

par(mfrow = c(1,2))

hist(migraine$NDDIE[which(migraine$Migraine == 1)], xlim = c(0,50), col = 'red', xlab = 'NDDIE', main = 'With Migraine')
hist(migraine$NDDIE[which(migraine$Migraine == 0)], xlim = c(0,50), col = 'cyan', xlab = 'NDDIE', main = 'Without Migraine')
hist(migraine$CESD[which(migraine$Migraine == 1)], xlim = c(0,50), col = 'red', xlab = 'CSED', main = 'With Migraine')
hist(migraine$CESD[which(migraine$Migraine == 0)], xlim = c(0,50), col = 'cyan', xlab = 'CSED', main = 'Without Migraine')
hist(migraine$`ABNAS memory`[which(migraine$Migraine == 1)], xlim = c(0,50), col = 'red', xlab = 'ABNAS MEMORY', main = 'With Migraine')
hist(migraine$`ABNAS memory`[which(migraine$Migraine == 0)], xlim = c(0,50), col = 'cyan', xlab = 'ABNAS MEMORY', main = 'Without Migraine')
hist(migraine$`ABNAS language`[which(migraine$Migraine == 1)], xlim = c(0,50), col = 'red', xlab = 'ABNAS LANGUAGE', main = 'With Migraine')
hist(migraine$`ABNAS language`[which(migraine$Migraine == 0)], xlim = c(0,50), col = 'cyan', xlab = 'ABNAS LANGUAGE', main = 'Without Migraine')

```


**We do the similar process as explained above using boxplots.**

```{r boxplot}

layout(heights = c(5,5),T)
par(mfrow = c(1,2))

boxplot(migraine$NDDIE[which(migraine$Migraine == 1)], col = 'red', xlab = 'NDDIE', main = 'With Migraine')
boxplot(migraine$NDDIE[which(migraine$Migraine == 0)], col = 'cyan', xlab = 'NDDIE', main = 'Without Migraine')
boxplot(migraine$CESD[which(migraine$Migraine == 1)], col = 'red', xlab = 'CSED', main = 'With Migraine')
boxplot(migraine$CESD[which(migraine$Migraine == 0)], col = 'cyan', xlab = 'CSED', main = 'Without Migraine')
boxplot(migraine$`ABNAS memory`[which(migraine$Migraine == 1)], col = 'red', xlab = 'ABNAS MEMORY', main = 'With Migraine')
boxplot(migraine$`ABNAS memory`[which(migraine$Migraine == 0)], col = 'cyan', xlab = 'ABNAS MEMORY', main = 'Without Migraine')
boxplot(migraine$`ABNAS language`[which(migraine$Migraine == 1)], col = 'red', xlab = 'ABNAS LANGUAGE', main = 'With Migraine')
boxplot(migraine$`ABNAS language`[which(migraine$Migraine == 0)], col = 'cyan', xlab = 'ABNAS LANGUAGE', main = 'Without Migraine')

```



